\chapter*{Acknowledgments}

At first, I want to thank my advisor Prof. Rhode from whom I have learned many things since I started my Bachelor thesis at his chair in 2014.
He gave me the opportunity to do this work and invested a lot of time advising me and discussing problems.
Thank You for your support.
Furthermore, I want to thank Prof Spaan, who agreed to be the second assessor of this thesis.

I also want to thank the chair of Prof. Rhode, who welcomed me warmly when I started as a Bachelor, and ever since I experienced a supporting and kind atmosphere towards me.
Especially I want to thank Alexander Sandrockwith my deepest gratitude, who invested a lot of time in me and thereby had a lot of patience. But we had a good collaboration and I still appreciate working together with you.
Furthermore, I want to thank my office, Thorben Menne, Maximilian Meier, and Mario Dunsch for the good working atmosphere. I appreciate it, especially in the context of writing this acknowledgment after more than one year of the pandemic.
I want to highlight Mirco Hünnefeld and Mathis Börner and thank you for the numerous discussions.
To the PROPOSAL development squad, including Mario Dunsch, Jean-Marco Alameddine, and Maximilian Sackel, Alexander Sandrock, Maximilian Meier, and Thorben Menne I just want to say, that you are the best.
Special thanks thereby to Maximilian Nöthe for his special help.
Finally, I want to thank the heart of this chair, thank you, Andrea Teichmann.

I acknowledge the DFG for funding this muon project, the Wilhelm and Else Heraeus Foundation for funding the DPG conferences, and the DAAD for the 3-month scholarship.
Thereby, I want to thank Prof. Claudio Kooper and the whole IceCube group at the Michigan State University for the interesting discussions and insights I could get.
Huge thanks also to the complete IceCube collaboration for the fruitful discussions. I gained a lot of experience and it was an honor to work in one of the most exciting experiments on earth.
I also want to thank the CORSIKA collaboration for their support, their ideas, and just the very good collaboration.

Finally, I thank my family.
No words can describe how much you have always supported me and how I can trust in you, which gives me strength.
Thank You